\chapter{Agentspark}
\label{cha:agentspark}

This chapter introduces AgentSpark, the demo agent implementation
available for SimSpark.

\section{Behaviors}

AgentSpark is a demo agent implementation for use with the SimSpark
server. It serves as a testbed for agent implementation. It implements
all low level details of connecting and communicating with the
server. It further implements an abstract run loop. The behavior of
this agent skeleton is implemented using plugins. On of these plugins
is configured and installed at run time.

\subsection{SoccerbotBehavior}

The \texttt{SoccerbotBehavior} is a minimal example of an agent that
acts on a soccer field and controls our current humanoid soccer bot
model. It demonstrates reading preceptor values and controlling the
robot with the installed effectors in a control loop. The implemented
behavior resembles the classic \texttt{hello world} program in a
way. It moves up the arms of the robot and repeatedly waves hello.

\section{How to change Behaviors?}

The different Behaviors of AgentSpark are encapsulated in classes
that derive from the \texttt{Behavior} class. Classes of this type
implement an interface with two functions that each return a command
string that is then send to the SimSpark server.

The first implemented function is \texttt{Init}. This function is
called once when the agent initially connected to the server. It is
typically used to construct the agent representation in the server
with the help of the \texttt{scene} effector and to move the agent to
a suitable start position in the simulation.

The second implemented function is called \texttt{Think}. This
function is called every simulation cycle. As a parameter it receives
the sensor data as reported from the server in the last cycle. This
function should implements the main behavior run loop of each
agent.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "user-manual"
%%% End: 
