\chapter{Introduction}

\section{What is SimSpark?}

SimSpark is a multi-agent simulation system for agents in
three-dimensional environments. Its goal is to provide a high degree
of flexibility for creating new types of simulations. It builds on a
flexible application framework and exhausts the idea of replaceable
components throughout its implementation.

In comparison to specialized simulators, users can create new
simulations by using a scene description language. SimSpark is a
powerful tool to state different multi-agent research questions and is
used as the official simulator for the RoboCup Simulation League
competition.

\section{History of the Project}

SimSpark was developed as part of the RoboCup initiative, initially
called Robot World Cup Initiative. It is an international research and
education initiative. It is an attempt to foster AI and intelligent
robotics research by providing a standard problem where wide range of
technologies can be integrated and examined, as well as being used for
integrated project-oriented education.

For this purpose, RoboCup chose to use soccer game as a primary
domain, and organizes competitive RoboCups. The simulation league is
one of several leagues where the entire soccer game takes places in a
simulated environment. SimSpark's two dimensional predecessor
simulation models the players and the ball as flat spheres. It further
lacks a realistic physical environment.

As one of the long term goals of the soccer simulation is to aim for
realism the long term objective are realistic humanoid players in a
physical environment. These players should one day challenge the
champion of the most recent World Cup.

Therefore, on the RoboCup 2003 Symposium a new approach to a
three-dimensional physically realistic soccer simulation was
proposed~\cite{KO04}. In a road map discussion for the Soccer Simulation League on
RoboCup 2003, the participants decided on adding the three-dimensional
simulation to the competitions.

For RoboCup 2004, SimSpark was successfully used for the first
official competition in RoboCup Simulation League 3D.  The soccer
simulation for this tournament was developed in parallel with the
SimSpark simulator. In its initial version players were modeled as
spheres in a physical three dimensional world. Since then SimSpark
grew considerably and now supports humanoid players with
articulated bodies.

It served from the beginning as a test bed and a guide for essential
new features that were added to the simulator during
development. However changes to the simulator core were never
customized for the soccer simulation. Instead generic
simulator services were implemented with all soccer specific
details contained in a set of plugins (see \cite{OR05}).

\section{About this Manual}

% add contact and download location

This manual describes the SimSpark simulator. Like the simulator
itself it is subject to constant change in an ongoing development
effort. It assumes that you are familiar with the basic concepts of
multiagent simulations.

It aims to be a guide on how to develop your own RoboCup agents,
construct new robot models, build your own custom monitor or use the
trainer command protocol to test your agents.

If there are errors, inconsistencies, or oddities, please notify
<EMail Address> with the location of the error and a suggestion of how
it should be corrected. We are always looking for anyone who has an
idea on how to improve the manual, as well as proofread or rewrite a
section of the manual.

The latest manual can be downloaded at <URL>.

\section{Reader's Guide to the Manual}

% \ref{cha:overview}

% short introduction to the system and its components; 
% software architecture view:different libraries take take care of different aspects of the simulator
% component view: server, monitor, logplayer, agent, available simulations and robot models


% \ref{cha:getstarted}
% helps to get the user started; guide through a typcial install process; 
% gives an example simulation run


% \ref{cha:simspark}
% detailed information about the messages that are sent from the
% server to the agent (perceptors) and vice versa (effectors)


% \ref{cha:monitorspark}
% an overview about the available monitor and log
% player setups available  their configuration and usage


% \ref{cha:agentspark}
% explain demo agent; how to configure different behaviors


% \ref{cha:simulations}
% introduce available simulations for SimSpark; describe soccer simulation in more detail


% \ref{cha:robots}
% describe available robot models that are available for SimSpark; describe the currenr SoccerBot in more detail


% \ref{cha:resources}
% where to download the server, how to contact the developer community and report bugs; 
% links to papers and thesis works related to SimSpark

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "user-manual"
%%% End: 
