\chapter{Rcssmonitor3d}
\label{cha:rcssmonitor3d}

This chapter gives an overview about the available monitor and log
player setups available in SimSpark, their configuration and usage.

\section{Internal Monitor}

The internal monitor implementation is part of the SimSpark server. It
is enabled in the \texttt{simspark.rb} setup script by enabling the
rendering and input plugins of the server. To do so please uncomment
the lines \texttt{sparkSetupRendering()} and
\texttt{sparkSetupInput()}, i.e. remove the leading \texttt{'\#'} comment markers.

\section{External Monitor}

The external monitor implementation is called
\texttt{rcssmonitor3d}. It either connects to a running SimSpark
instance or replays a simulation run from a log file.

\section{Playing Log files}

The monitor is able to replay a simulation that is recorded in a log
file. It accepts a \texttt{--logfile} parameter on its command line
that enables the log player mode.

\section{Key Assignments}

The monitor accepts a number of commands by key or mouse. These
commands either control the movement of the camera or send commands
back to the server to control the simulation. The soccer simulation
uses these commands to implement the controls for the human referee.

\begin{table}
\label {table:rcssmonitor3d}
\caption{Key assignments for rcssmonitor3d}
\begin{center}
\begin{tabular}{|l|l|}
{\bf key} & {\bf function} \\
\hline
q & quit monitor \\
\hline
left mouse button & mouse look \\
right mouse button & move camera up \\
keypad plus & move camera up \\
pageup & move camera up \\
pagedown & move camera down \\
keypad minus & move camera down \\
a & move camera left \\
left arrow & move camera left \\
d & move camera right \\
right arrow & move camera right \\
w & move camera forward \\
up arrow & move camera forward \\
s & move camera backward \\
down arrow & move camera backward \\
\hline
1 & camera to left goal \\
2 & camera to left corner \\
3 & camera to middle left \\
4 & camera to middle right \\
5 & camera to middle \\
6 & camera to right corner \\
7 & camera to right goal \\
\hline
l & free kick left \\
r & free kick right \\
k & kick off \\
b & drop ball \\
m & move agent \\
n & shoot ball \\
x & move ball \\
\hline
p & pause the playback of a log file \\
f & move forward in the log file \\
b & move backwards in the log file \\
l & playback the log file \\
\hline
\end{tabular}
\end{center}
\end{table}

The list of accepted default command short cuts is given in table
\ref{table:rcssmonitor3d}. These short cuts are configured in the two setup
scripts \texttt{bindings.rb} and \texttt{soccerbindings.rb} that the
SimSpark monitor reads on start up.

 %%% Local Variables: %%% mode: latex %%% TeX-master:
"user-manual" %%% End:
